### 2009-01-28 trunk/Papers/rmrs-08/entailment.tex
EP fixed
final version?

The above definitions serve an important theoretical purpose in that
they provide a formal underpinning for the use of \rmrs\ in practical
systems.  Next to the peace of mind that comes with the use of a
well-understood formalism, we hope that the work reported here will
serve as a starting point for future research.

One direction which we believe may develop from this paper is the
development of efficient solvers for \rmrs.  As a first step, it would
be interesting to define a practically useful fragment of \rmrs\ with
polynomial-time satisfiability.  Our definition is sufficiently close
to that of dominance constraints that we expect that it should be
feasible to carry over the definition of \emph{normal dominance
  constraints} \cite{Althaus_etal:JoA} to \rmrs; neither the lexical
ambiguity of the node labels nor the separate specification of
predicates and arguments should make satisfiability harder.

@@

At this point, we cannot provide an efficient algorithm for testing
entailment of \rmrs.  However, we propose the following novel
syntactic characterisation as a starting point for research along
those lines. We call an \rmrs\ $\varphi'$ an \emph{extension} of the
\rmrs\ $\varphi$ if $\varphi'$ contains all the EPs of $\varphi$ and
$D(\varphi') \supseteq D(\varphi)$.

>>>>

The above definitions serve an important theoretical purpose: they
formally underpin the use of \rmrs\ in practical
systems.  Next to the peace of mind that comes with the use of a
well-understood formalism, we hope that the work reported here will
serve as a starting point for future research.

One direction to pursue from this paper is the
development of efficient solvers for \rmrs.  As a first step, it would
be interesting to define a practically useful fragment of \rmrs\ with
polynomial-time satisfiability.  Our definition is sufficiently close
to that of dominance constraints that we expect that it should be
feasible to carry over the definition of \emph{normal dominance
  constraints} \cite{Althaus_etal:JoA} to \rmrs; neither the lexical
ambiguity of the node labels nor the separate specification of
predicates and arguments should make satisfiability harder.

@@

At this point, we cannot provide an efficient algorithm for testing
entailment of \rmrs.  However, we propose the following novel
syntactic characterisation as a starting point for research along
those lines. We call an \rmrs\ $\varphi'$ an \emph{extension} of the
\rmrs\ $\varphi$ if $\varphi'$ contains all the {\sc ep}s of $\varphi$ and
$D(\varphi') \supseteq D(\varphi)$.

### 2009-01-28 trunk/Papers/rmrs-08/rmrs.tex
final version?

The {\em satisfaction} relation between an \rmrs\ $\varphi$ and a
finite constructor tree $\tau$ is defined in terms of several
assignment functions.  First, a node variable assignment
function $\alpha: \nvar \rightarrow D(\tau)$ maps the node
variables in an \rmrs\ to the nodes of $\tau$.  Second, a
base language assignment function $g:\ovar \rightarrow \Sigma_0$ 
maps the base variables to nullary constructors
representing variables in the base language. Finally, a
relation $\sigma \subseteq \pred \times \Sigma_{\geq 1}$ maps
each \rmrs\ predicate symbol to a set of constructors from $\Sigma$.
As we'll see shortly, this relation allows an \rmrs\ to underspecify lexical
ambiguities.

@@

So far we have based the syntax and semantics of \rmrs\ on the
dominance relation from \newcite{egg:etal:2001} rather than the qeq
relation from \newcite{copestake:etal:2005}.  This is partly because
dominance is a weaker relation than qeq: If a dependency parser links
a determiner to a noun and this noun to a verb, then we can use
dominance but not qeq to represent that the predicate introduced by
the verb is outscoped by the quantifier introduced by the determiner
(see earlier discussion).  However, it is very straightforward to add
qeq to the syntax of the language and to extend the model theory
accordingly.  This extension adds a new atom $X \qeq Y$ to
Def.~\ref{defn:rmrs-syntax}, and $\tau,\alpha,g,\sigma$ will satisfy
$X \qeq Y$ iff $\alpha(X) \dom \alpha(Y)$, each node on the path is a
quantifier, and each step in the path goes to the rightmost child. All
the above propositions about solved forms still hold if ``dominance''
is replaced with ``qeq''.

>>>>

The {\em satisfaction} relation between an \rmrs\ $\varphi$ and a
finite constructor tree $\tau$ is defined in terms of several
assignment functions.  First, a node variable assignment
function $\alpha: \nvar \rightarrow D(\tau)$ maps the node
variables in an \rmrs\ to the nodes of $\tau$.  Second, a
base language assignment function $g:\ovar \rightarrow \Sigma_0$ 
maps the base variables to nullary constructors
representing variables in the base language. Finally, a
function $\sigma$ from $\pred$ to the power set of $\Sigma_{\geq 1}$ maps
each \rmrs\ predicate symbol to a set of constructors from $\Sigma$.
As we'll see shortly, this relation allows an \rmrs\ to underspecify lexical
ambiguities.

@@

So far we have based the syntax and semantics of \rmrs\ on the
dominance relation from \newcite{egg:etal:2001} rather than the qeq
relation from \newcite{copestake:etal:2005}.  This is partly because
dominance is the weaker relation: If a dependency parser links
a determiner to a noun and this noun to a verb, then we can use
dominance but not qeq to represent that the predicate introduced by
the verb is outscoped by the quantifier introduced by the determiner
(see earlier discussion).  However, it is very straightforward to
extend the syntax and semantics
of the language to include the qeq relation.
This extension adds a new atom $X \qeq Y$ to
Def.~\ref{defn:rmrs-syntax}, and $\tau,\alpha,g,\sigma$ will satisfy
$X \qeq Y$ iff $\alpha(X) \dom \alpha(Y)$, each node on the path is a
quantifier, and each step in the path goes to the rightmost child. All
the above propositions about solved forms still hold if ``dominance''
is replaced with ``qeq''.
